---
title: "Calculation of altitude (zonal statistics)"
author: "Kadlec Ivo"
date: "2024-06-10"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_float: yes
    highlight: pygments
    theme: flatly
    number_sections: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=200, echo = TRUE, message = FALSE, warning = FALSE)
```

```{r library, warning=FALSE}
library(sf)
library(elevatr)
library(raster)
library(dplyr)
library(tidyverse)
```

# Loading data
Data is loaded in csv format, which must include coordinates (it is recommended to have them named "X" and "Y"). Other information/columns are optional and included in the final dataset with elevation. 

```{r, message=FALSE}
data <- read.csv("data.csv", sep = ";", header = T, dec = ",") # load data
```

# Preparation for analysis
The script uses the "elevatr" package to retrieve a raster layer containing elevation data, which allows access to web services and returns elevation data in the form of either points ("sf" object) or a raster ("raster" object). By default, the script uses AmazonWeb Services. 
```{r, warning=FALSE}
# Converting coordinates to points and then creating a buffer around them
# Note: The size of the buffer is specified in meters in this case, it may vary if using a different coordinate system.
df_sf <- sf::st_as_sf(data, coords = c("X", "Y"), crs = 4326)
df_sf_buf <- st_buffer(df_sf, 8000) 

# Loading the raster
elevation <- get_elev_raster(df_sf_buf, z = 11) #import elevation raster

# Preview of the raster with altitude and the created polygons
plot(elevation)
plot(st_geometry(df_sf_buf), add = TRUE, col = "blue")
```

# Zonal statistic
```{r}
# Zonal statistics
df_alt <- raster::extract(elevation,df_sf_buf, fun=mean) %>% 
  cbind(data,.) # Linking the results to the original dataset
names(df_alt)[ncol(df_alt)] <- "altitude [m]"
head(df_alt)

write.csv(df_alt, "pack_alt_DEU.csv", sep=";") # Save as csv
```

